\documentclass[onecolumn]{aastex62}
\usepackage{times}
\usepackage{epsfig}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{color}
\usepackage{microtype}
\usepackage{float}
\usepackage{stfloats}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{placeins}
\graphicspath{{./figures/}}
\usepackage{wrapfig}
\hypersetup{backref,breaklinks,colorlinks,citecolor=blue}
\usepackage[all]{hypcap}
\usepackage{definitions}
\usepackage{subfiles}

\bibliographystyle{apj}

\newcommand{\subdate}{\today}
\newcommand{\shortauth}{Barker et al.}
\newcommand{\slugcom}{Draft - \today}
\newcommand{\Pos}{\textnormal{\tiny\textsc{Pos}}}
\newcommand{\TVD}{\textnormal{\tiny\textsc{Tvd}}}
%\newcommand{\slugcom}{Submitted to ApJL on \subdate}
%\slugcomment{\slugcom}


\lefthead{\sc \footnotesize \slugcom \hfill \shortauth}
\righthead{\sc \footnotesize \slugcom \hfill \shortauth}
%}

\bibliographystyle{aasjournal}
\received{\today}
\shorttitle{Discontinuous Galerkin methods with a nuclear EOS}

\begin{document}

\title{\thornado-hydro: Generalizing discontinuous galerkin methods for a nuclear
equation of state for supernova hydrodynamics}
\author{Brandon Barker}
\affiliation{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996}

\author{Eirik Endeve}
\affiliation{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996}
\affiliation{Computer Science and Mathematics Division, Oak Ridge National Laboratory, TN 37831}
\affiliation{Joint Institute for Computational Sciences, Oak Ridge National Laboratory, TN 37831}

\author{Anthony Mezzacappa}
\affiliation{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996}
\affiliation{Joint Institute for Computational Sciences, Oak Ridge National Laboratory, TN 37831}

%\begin{abstract}
%\end{abstract}
% nonlinear hyperbloic cons. laws. develop discontinuities from smooth ICs
% this can lead to oscillatins which require some limiting process to smooth.
% limiting in characteristic variables increases the efficiency of slope limiting
\begin{abstract}
A problem of high importance in computational astrophysics is obtaining accurate
solutions to the Euler equations of hydrodynamics. We are interested in solving
the Euler equations in the context of core collapse supernovae.
The toolkit for high-order neutrino-radiation hydrodynamics (\thornado)
is being developed for CCSNe simulations and related problems utilizing a spatial discretization
based on the discontinuous Galerkin (DG) method.
The Euler equations form a hyperbolic set of partial differential equations.
In the quasi-linear form, the system can be represented
as a set of independent advection equations that can be limited separately.
This use of characteristic variables also increases the efficiency of the slope
limiting process. However, the addition of the nuclear equation of state equation
to the Euler equations makes the decomposition nontrivial.
We introduce the framework for the characteristic decomposition of the Euler equations
with the inclusion of the nuclear EOS terms and
present results from some initial tests.
\end{abstract}

\keywords{supernovae: general -- hydrodynamics -- convection -- turbulence -- equation of state -- methods: numerical -- stars: massive -- stars: evolution}


\section{Introduction}
\label{sec:Intro}
\todo{Fill in introduction. Contextualize the work.}
The core collapse supernovae (CCSNe) explosion mechanism is fundamentally three-dimensional
in nature \citep[see e.g.,][]{blondin:2006, muller:2012, oconnor:2018b}. Alongside general
reletavistic gravity, complex nuclear equations of state (EOS), and
neutrino transport, hydrodynamics must be accurately modelled,
creating a very challenging and compelling problem. Hydrodynamic instabilities are critical
in aiding the explosion, with complex phenomena such as turbulence and
convection playing key roles in the CCSN mechanism
\citep{murphy:2011, murphy:2013, couch:2013, couch:2015a, radice:2016, mabanta:2018}.
The accurate and efficient modelling of the supernova hydrodynamics is crucial if the
explosion is to be realistically modelled.
For in-depth reviews of the CCSN mechanism, see \citet{bethe:1990, janka:2007, janka:2012a, janka:2016, burrows:2013, hix:2014, muller:2016, couch:2017}.

\todo{Talk about \thornado, DG methods, Euler Equations, possibly reference other codes that
employ different methods and those methods' drawbacks.}

This paper will be organized as follows: in Section \ref{sec:eulereq} we briefly
describe the Euler equations of gas dynamics. Section \ref{sec:DG} describes our
numerical implementation of the DG method in the one dimensional (1D) case.
Section \ref{sec:results} shows the results of 1D hydrodynamic simulations,
and Section \ref{sec:eigen} presents the eigenvectors used in the eigen-decomposition
of the Euler equations.

\section{Euler Equations of Gas Dynamics in Cartesian Coordinates}
\label{sec:eulereq}
\todo{Use this form of the Euler Equations?}
The non-relativistic Euler equations of gas dynamics
\citep[see, e.g.,][for details]{leveque:2002} in Cartesian coordinates
in the absence of sources with a nuclear matter EOS are given by
the equations of conservation of mass,
\beq
  \partial_{t} \rho + \divergence{} (\rho\,  \mathbf{v}) = 0
  \label{eq:massConservation}
\eeq
conservation of momentum,
\beq
%	\partial_{t}(\rho \mathbf{v}) + \divergence{}(\rho u^2 + p) = 0
  \partial_{t}(\rho\,  \mathbf{v}) + \divergence{}(\rho\,  \mathbf{v} \otimes \mathbf{v} + P\, \mathbf{I}) = 0
  \label{eq:momentumConservation}
\eeq
conservation of energy,
\beq
  \partial_{t} E + \divergence{}\left[(E+P)\, \mathbf{v}\right] = 0
  \label{eq:energyConservation}
\eeq
and conservation of \textcolor{red}{electron density}
\beq
  \partial_{t} D + \divergence{}(D\, \mathbf{v}) = 0
  \label{eq:electronConservation}
\eeq
where $\rho$ represents mass density, $\mathbf{v}$ the fluid velocity vector,
$P$ the fluid pressure, $D=\rho y_e$ where $y_e$ is the electron fraction,
$E=\epsilon \rho +\frac{1}{2}\rho v^2$ the total energy (internal plus kinetic),
$\epsilon$ is the specific internal energy, and $\mathbf{I}$ is the identity tensor.
\eqref{eq:massConservation}-\eqref{eq:electronConservation}. The inclusion of
Equation \eqref{eq:electronConservation} is because we require a nuclear EOS.
These equations are closed by a tabulated EOS where the pressure is given by a function of
density, temperature $T$ and the electron fraction: $P = P(\rho, T, y_e)$.
We may rewrite Equations \eqref{eq:massConservation}-\eqref{eq:electronConservation}
in a more convenient way:

\beq
  \partial_{t}\mathbf{U} + \partial_{x}\mathbf{F}(\mathbf{U}) = 0,
  \label{eq:conservation}
\eeq
where $\mathbf{U} =(\rho,\rho \mathbf{v},E, D)^{T}$ is the vector of conserved quantities
and $\mathbf{F}(\mathbf{U})=(\rho \mathbf{v},\rho u^{2}+p,(E+P)\mathbf{v}, D\mathbf{v})^{T}$
is the flux vector.

\section{Numerical Implementation}
\label{sec:DG}

\todo{Add intro paragraph} \\
\textcolor{red}{Subsections or no?}

\subsection{The Discontinuous Galerkin Methods}
In our solver we have chosen the discontinuous Galerkin (DG) method
for our spatial discretization. In this section we will briefly discuss our
implementation of the DG method, introducing notation and concepts.
For simplicity, we will focus on the one dimensional (1D) case. Recall that
we seek solutions to the Euler equations of hydrodynamics, constituting a
conservation law of the form

\beq
  \partial_{t} \mathbf{U} + \partial_{x} \mathbf{F}(\mathbf{U}) = 0
  \label{eq:ode}
\eeq
Where $\mathbf{U}$ is the evolved state vector and $\mathbf{F}(\mathbf{U})$ is
the flux. In order to solve Equation \eqref{eq:ode} numerically, we divide the computational
domain $D\subset \mathbb{R}$ into a disjoint union $\mathcal{T}$ of open elements
$\bK$ such that $D = \cup_{\bK \in \cT}\bK$. Each element $\bK$ is a box in the
coordinates
\beq
  \bK \in (\xL,x_{R}),
\eeq
We let the approximation space $\mathbb{V}^{k}$ for the DG method
be constructed from the tensor product of one-dimensional
polynomials of degree $k$. Note that functions in $\mathbb{V}^{k}$
can be discontinuous across element interfaces. The DG problem is then to find
$\mathbf{U}_h \in \mathbb{V}^{k}$ which approximates $\mathbf{U}$ in Equation
\eqref{eq:ode}, such that $\forall$ $\phi \in \mathbb{V}^{k}$ and $\bK \in \cT$

\beq
  \partial_{t} \int_{\bK}\mathbf{U}_h\, \phi\, dx +
  \int_{\bK}\partial_{x}\, \mathbf{F(\mathbf{U})}\, \phi\, dx = 0.
\label{eq:semidiscreteDG_almost}
\eeq
Integrating the second term by parts, this becomes
\beq
   \partial_{t} \int_{\bK}\mathbf{U_{h}}\, \phi\, dx +
   \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{-} \big|_{x_{R}}
   + \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{+} \big|_{x_{L}} +
   \int_{\bK}\mathbf{F(\mathbf{U}_h)}\, \partial_{x}\phi\, dx = 0.
\label{eq:semidiscreteDG}
\eeq
In Eq.~\eqref{eq:semidiscreteDG}, $\widehat{\vect{F}}(\vect{U}_{h})$ is a
numerical flux approximating the flux on the boundary of $\bK$.
The numerical flux function is evaluated using values from
both boundaries of an element; i.e.,
\begin{equation}
  \widehat{\vect{F}}(\vect{U}_{h})=\vect{f}(\vect{U}_{h}(x^{-}),\vect{U}_{h}(x^{+})),
\end{equation}
where superscripts $-/+$
indicate that the function is evaluated to the immediate left/right of the
interface. We use the Harten-Lax-van Leer (HLL) flux \citep{harten:1983}
for all the numerical experiments presented in Section \ref{sec:results}.

In each element $K$, we use a nodal representation of the conserved variables
$\mathbf{U}$:

\beq
\begin{split}
  \mathbf{U}(x,t) \approx\, & \mathbf{U}_h (x,t) =
  \sum_{i=1}^{N}\vect{U}_{i}(t)\,\ell_{i}(x),
  \quad\text{where}\quad \\
  & \ell_{i}(\eta)=
  \prod_{\substack{j=1\\j\ne i}}^{N}\f{\eta-\eta_{j}}{\eta_{i}-\eta_{j}}
  \label{eq:conservedNodalExpansion}
\end{split}
\eeq
are Lagrange polynomials defined on $I = \{ \eta : \eta \in (-0.5,0.5) \}$,
and are constructed to interpolate the node set
$S_{N}=\{\eta_{i}\}_{i=1}^{N}\subset I$. The spatial coordinate $x$ and the
reference coordinate $\eta$ are related by the
mapping $x(\eta)=\xL+(0.5+\eta)\,\Delta x$.
Then, for any $\eta_{j}\in S_{N}$, $\ell_{i}(\eta_{j})=\delta_{ij}$,
so that $\vect{U}_{h}(x(\eta_{j}),t)=\vect{U}_{j}(t)$.
First we define the $M$-point quadrature $Q_{M}:C^{0}(I)\to\mathbb{R}$
with abscissas $\hat{S}_{M}=\{\eta_{q}\}_{q=1}^{M}$ and weights
$\{w_{q}\}_{q=1}^{M}$, normalized such that $\sum_{q=1}^{M}w_{q}=1$ in order to
evaluate the integrals in Eq.~\eqref{eq:semidiscreteDG}.
We use the $M$-point Legendre-Gauss quadrature, which is exact for
polynomials of degree $\le 2M-1$.
Then, if $P_{h}(x)$ is such a polynomial, we have
\beq
  \f{1}{\Delta x}\int_{K}P_{h}(x)\,dx=\int_{I}P_{h}(\eta)\,d\eta=\sum_{q=1}^{M}w_{q}\,P_{h}(\eta_{q}).
\eeq

For the sake of efficiency we let $M=N$ and $S_{N}=\hat{S}_{N}$, which is a
spectral-type nodal collocation DG approximation \citep{bassi:2013}.
Inserting Eq.~\eqref{eq:conservedNodalExpansion} into Eq.~\eqref{eq:semidiscreteDG},
letting $\phi(x)=\ell_{k}(x)$, using the quadratures defined above,
we obtain
\begin{align}
  \pd{}{t}\int_{K}\vect{U}_{h}\,\phi\,dx
  &\approx w_{k}\,\pd{}{t}\vect{U}_{k}\,\Delta x
  \label{eq:timeDerivativeTerm}
\end{align}
for the time derivative piece.
Similarly, the last term on the left-hand side of Eq.~\eqref{eq:semidiscreteDG}
becomes
\beq
  \int_{K}\vect{F}(\vect{U}_{h})\,\pderiv{\phi}{x}\,dx
  \approx \sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q}).
  \label{eq:volumeTerm}
\eeq
Now we may combine Eqs.~\eqref{eq:timeDerivativeTerm}-\eqref{eq:volumeTerm} in
Eq.~\eqref{eq:semidiscreteDG} resulting in the semi-discrete form
\beq
  \deriv{\vect{U}_{k}}{t}
   =-\f{1}{w_{k}\Delta x}
  \Big\{
  \Big[\,
    \widehat{\vect{F}}\ell_{k}\big|_{x_{H}}
     -\widehat{\vect{F}}\ell_{k}\big|_{\xL}
  \,\Big]
   -\sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q})
  \Big\}.
  \label{eq:semidiscreteDiscretized}
\eeq
We now have a system of ordinary
differential equations (ODEs), which may be evolved in time with an ODE solver.
In Section~\ref{sec:results} we use the third-order strong
stability-preserving Runge-Kutta (SSP-RK3) method \citep{shu:1988}.

\subsection{Time Integration}
A common feature of high-order numerical PDEs is unphysical oscillations in the solutions.
It is therfore of great interest in the DG algorithm to implement slope limiting
of the polynomial $\mathbf{U}_h$. We use the total variation diminishing (TVD)
slope limiter \citep[see, e.g.,][]{cockburn:1998} in conjunction with the
troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting
by flagged elements where limiting is needed.
The general $s$-stage Runge-Kutta time stepping algorithm, including
the limiting process, can then be summarized as in \citet{cockburn:2001}:
\begin{itemize}
  \item[1.] Set $\bar{\vect{U}}^{(0)} = \bar{\vect{U}}^{n}$,
  \item[2.] For $i=1,\ldots,s$ compute:
  \begin{equation}
  \small  \bar{\vect{U}}^{(i)}
    = \Big\{\Lambda^{\TVD}\Big\{\sum_{j=0}^{i-1}\alpha_{ij}\,\bar{\vect{U}}^{(j)}+\beta_{ij}\,\Delta t\,\bar{\vect{F}}\big(\bar{\vect{U}}^{(j)}\big)\Big\}\Big\},
    \label{eq:rkStages}
  \end{equation}
  \item[3.] Set $\bar{\vect{U}}^{n+1}=\bar{\vect{U}}^{(s)}$.
\end{itemize}
Above, the TVD limiter limiter preventing unphysical
states is denoted by the operators $\Lambda^{\TVD}\{\}$ The SSP-RK3 coefficients $\alpha_{ij}$
and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}. For each step
in Eq.~\eqref{eq:rkStages}, the TVD limiter is applied to elements flagged
by the troubled cell indicator. For a more in-depth discussion of the
limiting process used in \thornado, see \citet{endeve:2019}.

\subsection{Characteristic Decomposition}
Experience has shown that the slope limiting described in the previous section
is more effecient when performed on the so-called `characteristic variables`
as opposed to the conserved variables $\mathbf{U}_h$
\citep[see, e.g.,][for a description]{cockburn:1998}. We begin by
rewriting Eq.~\eqref{eq:conservation} in the quasi-linear form as follows
\beq
  \pderiv{\mathbf{U}}{t}
  + \pderiv{\mathbf{U}}{\mathbf{x}} \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}
  = 0.
  \label{eq:charEq}
\eeq
Because the Euler equations for a systemer of hyperblic
partial differential equations \citep[see, e.g.,][]{leveque:1992}, we can decompose the
Jacobian of the flux vector as

\beq
  \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}} =
  \mathcal{R} \Lambda \mathcal{R}^{-1},
\eeq
where the columns of $\mathcal{R}$ contain the right eigenvectors of the Jacobian,
the rows of $\mathcal{R}^{-1}$ contain the left eigenvectors, and
$\Lambda$ is a diagonal matrix containing the eigenvalues of the Jacobian.
At this point, we will introduce the characteristic variable
$\mathbf{w} = \mathcal{R}^{-1}\mathbf{U}$. Multiplying both sides of Equation
\eqref{eq:charEq} by $\mathcal{R}^{-1}$, we are able to linearize the system of equations to
a system of advection equations

\beq
  \pderiv{\mathbf{w}}{t} +
  \Lambda \pderiv{\mathbf{w}}{x}
  = 0.
\eeq
Solutions to these advection equations are far simpler to obtain than the
previous equations. Limiting may then be applied to dampen oscillations in the
solutions for the characteristic variables $\mathbf{w}$, which are then
transformed back to the solution of the conserved variables $\mathbf{U}$
\citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}. While this
process of characteristic limiting has been done for the ideal EOS
\citep[][]{cockburn:1998}, we want
to extend to the nuclear matter EOS case.


\section{Derivation of Jacobian and Eigenvalues for the 3D Nuclear Case}
\label{sec:eigen}

We assume that $P = P(\tau, \epsilon, n)$, where $\tau = \frac{1}{\rho}$.
Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D\}$,
where $m_i = \rho v_i$ and $D = \rho y_{e}$.
The flux vector is $\textbf{F}(\textbf{U}) =
\{m_{1}, m_{1}^{2}\tau + P, m_{1}m_{2}\tau, m_{1}m_{3}\tau,
(E+P)m_{1}\tau, Dm_{1}\tau\}$. The Jacobian of the flux vector is given by


%\begin{table*}[ht!]

\begin{align}
  \centering
	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
	= \left[
		\begin{array}{cccccc}
			0 & 1 & 0 & 0 & 0 & 0 \\
			-v_{1}^{2} -P_{\tau}\tau^{2} - P_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2}) & v_{1}(2-P_{\epsilon}\tau)  & -P_{\epsilon}v_{2}\tau & -P_{\epsilon}v_{3}\tau  & P_{\epsilon}\tau  & P_{D_{e}} \\
			-v_{1}v_{2} & v_2 & v_1 & 0 & 0 & 0 \\
			-v_{1}v_{3} & v_{3} & 0 & v_{1} & 0 & 0 \\
			v_{1}(-H - P_{\tau}\tau^{2} -P_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2})) & H - P_{\epsilon}v_{1}^{2}\tau  & -P_{\epsilon}v_{1}v_{2}\tau & -P_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+P_{\epsilon}\tau) & v_{1} P_{D_{e}} \\
			-v_{1} y_{e} & y_{e} & 0 & 0 & 0 & v_{1} \\
		\end{array}
    \right]
\end{align}
%\caption{}
%\end{table*}


\noindent Where $H=(E+P)\tau$ is the specific enthalpy and $P_{x} = \partial_{x}P$\footnote{Maybe define these elsewhere to show what's constant}. The eigenvalues of this Jacobian are given by

\begin{align}
\Lambda =
\begin{bmatrix}
  v_{1} - c_{s} & 0 & 0& 0& 0& 0 \\
  0 & v_{1} & 0 & 0 & 0 & 0      \\
  0 & 0 & v_{1} & 0 & 0 & 0      \\
  0 & 0 & 0 & v_{1} & 0 & 0      \\
  0 & 0 & 0 & 0 & v_{1} & 0      \\
  0 & 0 & 0 & 0 & 0 & v_{1} + c_{s}
\end{bmatrix}
\end{align}
where $c_{s} = \sqrt{\Gamma P \tau}$, for
$\Gamma = \frac{\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} y_{e} \tau^{-1}}{P}$, is
the local sound speed. In the ideal EOS regime this reduces to the sound speeds given \citet{colella:1985}.
The right eigenvectors are then given by
\begin{align*}
  \mathcal{R}_{1} =
  \left[
  \begin{array}{cccccc}
   1 & 0 & 1 & 1 & 0 & 1 \\
   v_{1}-c & 0 & v_{1} & v_{1} & 0 & c+v_{1} \\
   v_{2} & 1 & 0 & 0 & 0 & v_{2} \\
   v_{3} & 0 & 0 & 0 & 1 & v_{3} \\
   h-c v_{1} & v_{2} & \beta & 0 & v_{3} & h+c v_{1} \\
   y_{e}  & 0 & 0 & \frac{\tau  \chi }{2 P_{D_{e}}} & 0 & y_{e}  \\
  \end{array}
  \right]
\end{align*}
where the follow definitions have been used:
$h_{n} = \frac{c^2}{P_{\epsilon}\tau} + k $, $k = \frac{-y_{e} P_{D_{e}} \tau^{-1}
+ P_{\epsilon} (\frac{1}{2}v^2 + \epsilon) + P_{\tau}\tau}{P_{\epsilon}}$,
$\delta_{1} = v_{1}^{2}-v_{2}^{2}-v_{3}^{2}$,
$\chi = P_{\epsilon} ( \delta_{1} + 2\epsilon) + 2P_{\tau}\tau$, and
$\beta = \frac{1}{2} (\delta_{1}+2 \epsilon +\frac{2 P_{\tau} \tau }{P_{\epsilon}})$.
\begin{align*}
  \mathcal{R}_{1}^{-1} = \frac{1}{c^2}
  \left[
  \begin{array}{cccccc}
   \frac{1}{4} (2 c  v_{1}+\omega ) & \frac{1}{2} (-c- \phi_{1} ) & -\frac{1}{2} \phi_{2}
    & -\frac{1}{2} \phi_{3}  & \frac{P_{\epsilon} \tau }{2} & \frac{P_{D_{e}}}{2}
     \\
   -\frac{v_{2} \omega }{2} & \phi_{1} v_{2}  & c^2+\phi_{2} v_{2}  &
     \phi_{3} v_{2}  & -\phi_{2}  & -P_{D_{e}} v_{2}
     \\
   \frac{2 \chi  c^2+\alpha  \omega \tau^{-1} }{2 \chi } & -\frac{\phi_{1} \alpha  }{\chi \tau } &
     -\frac{\phi_{2} \alpha  }{\chi \tau } & -\frac{\phi_{3} \alpha }{\chi \tau } &
     \frac{P_{\epsilon} \alpha }{\chi } & \frac{P_{D_{e}} \left(\alpha -2 c^2\right)}{\tau \chi }
      \\
   -\frac{y_{e} P_{D_{e}} \omega }{\chi \tau } & \frac{2 y_{e} P_{D_{e}} \phi_{1} }{\chi \tau } & \frac{2 y_{e} P_{D_{e}}
     \phi_{2} }{\chi \tau} & \frac{2 y_{e} P_{D_{e}} \phi_{3} }{\chi \tau} & -\frac{2 y_{e}
     P_{D_{e}} P_{\epsilon} }{\chi } & \frac{2 P_{D_{e}} \left(c^2-y_{e} P_{D_{e}} \right)}{\tau  \chi }
      \\
   -\frac{v_{3} \omega }{2} & \phi_{1} v_{3}  & \phi_{2} v_{3}   &
     c^2+\phi_{3} v_{3}  & -\phi_{3}  & -P_{D_{e}} v_{3}
      \\
   \frac{1}{4} (\omega -2 c  v_{1}) & \frac{1}{2} (c-\phi_{1} ) & -\frac{1}{2} \phi_{2}
       & -\frac{1}{2} \phi_{3}  & \frac{P_{\epsilon} \tau }{2} & \frac{P_{D_{e}}}{2}
     \\
  \end{array}
  \right]
\end{align*}
where $\phi_{i} = P_{\epsilon}\,\tau\, v_{i}$,
$\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$, and
$\alpha = 2 y_{e} P_{D_{e}} - \tau \chi$.


\todo{What is the physical interpretation of h and k? I tried to follow \citet{schaal:2015a},
and in their eigenvectors (Ideal), h is specific enthalpy and k is specific kinetic energy.
What about my other defined variables? Should I redefined, or rename?}
The matrix of left eigenvectors is given by

\section{Preliminary Numerical Results}
\label{sec:results}
In this section we present preliminary numerical results obtained with the DG
method implemented in \thornado  leveraging use of the characteristic
decomposition. We use the SFHo EOS of nuclear matter \citep{steiner:2013}
now commonly used in hig-fidelity CCSN simulations.
These tests serve to gauge the performance of the
DG characteristic decomposition implementation on a set of benchmarks as an initial
assessment of its suitability for future CCSN simulations.

\subsection{Sod Shock Tube}
We present the classic 1D Sod shock tube Riemann problem \citep{sod:1978} in
Cartesian coordinates. The computational domain is D = [-5,5] km
with a discontinuity initially located at $x = 0$ km separating the left and right states

\beq
  \mathbf{U}_{L} = (10^{12} g/cm^3, 0\, km/s, 10^{32} ergs/cm^3, 0.4\rho)^T\,\,\,
  \mathbf{U}_{R} = (1.25*10^{11} g/cm^3, 0\, km/s, 10^{31} ergs/cm^3, 0.3\rho)^T
\eeq
\footnote{How to specify this? Specifically $y_e$}

The test is run until $t = 0.25$ms with 100 elements
\\
\todo{Componentwise vs Characteristic vs reference.}\\
Plots
\todo{Limiting Paramter Landscape.}\\
plots.
\todo{"Best" Parameters}\\
Plots
\todo{"Realistic" high density run}\\
Plots


\appendix

\section{Derivatives}
\input{derivatives}

\acknowledgements
We thank people.

\bibliography{ms}


\end{document}
